### 2015-06-28 abstract.tex
making changes...

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}

\noindent
Outline and writing assignments:
later the abstract will go here.
\begin{enumerate}
\item Introduction - Scott - Hasan - Gary - Jay 4 pages
   \begin{enumerate}
      \item resource drivers
      \item application drivers
      \item overall example which we will use for the test
      \item  need to answer the questions what we are doing, research  , prototype, aim, and what is new
   \end{enumerate}
\item Enabling Technologies - Jay - 1 page
   \begin{enumerate}
   	\item Sirocco
   	\item ADIOS
   	\item DataSpaces
   	\item Anything else
   \end{enumerate}
 \item {Proposed Research and Methods} - 14 pages
    \begin{enumerate}
       \item Application interface - Hasan, Manish, Gary, Kimmy, Jay, Mark
          \begin{enumerate}
            \item APIS - Hasan
            \item description of data - Gary
            \item collection of data reduction techniques - Mark
            \item data re-organization specification - Scott, Mark, Kimmy
            \item QoS - Hasan
            \item execution policies - Manish
            \item Initial placement and movement  - Gary
            \item policies - Jay
          \end{enumerate}
      \item Storage System - Jay, Carlos, Manish, Lee, Matthew, Feyi
      \begin{enumerate}
         \item migration - Lee
         \item resource management -Jay
         \item QoS - Carlos 
         \item Discovery - Jay
         \item Eviction metrics - Carlos
         \item Naming Service - Manish
         \item Estimation of Times - Feyi
      \end{enumerate}
    \end{enumerate}
\item Evaluation - 1 pages
\item Timetable of Activities - 1/2 page
\item Project Management Plan - 1 page
\item Project Objective  1/2 page
\end{enumerate}

>>>>

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}
Exascale scientific discovery will be
severely bottlenecked without sufficient new research into managing and
storing the large amounts of data that will be produced during the
simulation, and analyzed for months afterwards. In this project we will
demonstrate novel techniques to  facilitate efficient mapping of data
objects, even partitioning individual variables, from
the user space onto multiple storage tiers, and enable application-guided
data reductions/transformations to address capacity and bandwidth
bottlenecks. 
Our goal in this project is
to address the associated I/O and storage challenges in the context of
current and emerging storage landscapes, and expedite insights into mission
critical scientific processes; which is associated with {\bf theme two} of the FOA. To that end, we will build on the
capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
and services, the Sirocco peer-to-peer file system at
Sandia and the object storage and annotation expertise of UC Santa Cruz, to
explore application and multi-tier storage aware data management
solutions.    This project
brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

Our objective here is to reduce the time to knowledge, not just for a single
application, but for the entire workload in a multi-user environment, where
the storage is shared among users.
We achieve this goal by allowing selectable data quality, by trading its accuracy and error
in order to meet the time or resource constraint. 
We will explore beyond 
checkpoint/restart I/O, and will address the challenges posed by 
key
data access patterns in the knowledge gathering process.
Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
so that the best possible decisions can be autonomically made between the user intentions and
the available system resources.  
We will test our prototypes  on current and future DOE system with many of today's applications, including the
s XGC1, GTC, QMCPack, and SpecFM3D simulations. 
Our solutions will provide  new functionalities and APIs for
1) specifying, at the application level, data annotations that enable the
quantification of the relative importance and utility of data objects and
enable partitioning  across the storage hierarchy;
2) specifying selectable performance/quality/cost tradeoffs from both the
application and system perspectives;
3) evaluating these tradeoffs at runtime during data placement and movement,
and executing the resultant policies in an autonomic system using models,
heuristics and continuous learning; and
4) leveraging techniques such as application-aware data compression, 
and I/O prioritization to enforce these
policies.
\noindent

### 2015-06-28 narrative/intro.tex
making changes...

\begin{itemize}

\item describe data size problems for getting scientific results out
\item describe annotating and differentiating data storage according to both
  use and inherent data qualities (e.g., contains desireable features)
\item describe idea of using heavy compression for less interesting areas
  and less or no compression for more interesting areas.
\item describe inherent errors in scientific calculations
\item describe lossy compression within error bounds
\item describe how incorporating plug-ins for application aware data
  compression opportunities to aid data storage
\item describe Sirocco's current state and future designs as a solid
  foundation on which to work
\item describe how these new features can enhance a system like Sirocco to
  reduce data intensity
\item describe metadata challenges that we will also have to face.

\end{itemize}

every section should include:

Problem and background, one or more solution approaches, and investigation
goals/questions.

Technical Areas

1. Application Intentions.\\
2. Metadata Management (middleware and storage system)\\
3. Pluggable infrastructure for data transformation (storage system)\\
4. Intelligent (autonomic, application-aware) data mapping to storage/memory hierarchies (middleware and storage system)\\
5. Predictable Performance and resolution tradeoffs\\
6. Data re-generation. (middleware \& storage)\\
7. Learning Motifs and data re-organization.\\

\section*{Introduction}

The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project demonstrated that using a core
object storage system with integrated security providing authentication and
autorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has a usable storage infrastructure system. This
proposal seeks to build upon this work to demonstrate an application, data, and
usage aware storage system optimized for exascale composite applications and
capability application runs. We are proposing to combine the applications
knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
metadata and storage systems knowledge of UCSC, and the middleware expertise of
Rutgers to investigate how to make a data, workflow, and application aware
storage system.

>>>>

\section*{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
\label{sec:introduction}
The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, are added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project demonstrated that using a core
object storage system with integrated security providing authentication and
autorization services, customized storage interfaces could be developed with
exactly the overhead required for different applications. The second phase of
LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
devices for efficient and scalable storage performance.  Sirocco has been in
process for three years and has a usable storage infrastructure system. This
proposal seeks to build upon this work to demonstrate an application, data, and
usage aware storage system optimized for exascale composite applications and
capability application runs. We are proposing to combine the applications
knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
metadata and storage systems knowledge of UCSC, and the middleware expertise of
Rutgers to investigate how to make a data, workflow, and application aware
storage system.
